Dismal Code: Studying the Evolution of Security Bugs

### 2013-04-07 docs/evol.tex
more on the discussion

The main contributions of this research are:
\begin{itemize}
	\item The analysis of how security bugs evolve over time. To achieve
this, we inspect every project per version. Our hypothesis is that security
bugs must decrease as a project evolves since they are critical and developers
should eliminate them.
	\item Security bug persistence between versions. We show evidence that
security bugs are not eliminated between versions.
	\item We show evidence of the relation between security bugs and a project
version's size.  Here, size is defined in terms of bytecode.
	\item The correlation of security bugs with other bug categories. We
show evidence on how such bugs are correlated with bugs that have to do with
performance, coding practices and others.
%	\item The relation of software dependencies and security bugs. Our
%hypothesis is that an artifact that is a dependency of many other artifacts
%should have a low rate of security bugs. 
\end{itemize}

@@

Figure~\ref{fig:bug-per} shows how software bugs are distributed among the
repository. Together with the {\it Bad Practice} bugs and the {\it Style} bugs,
security bugs (the sum of the {\it Security} and {\it Malicious Code}
categories) are the most popular in the repository ($\geq 22\%$).
% An interesting observation is that 4,908 projects did not
% have any bugs at all. These cases involved small artifacts with a minimum number of classes.
% This goes along with the relationship between the size of an artifact and it's
% bugs that is presented in the upcoming section.

% dimitro@dds: maybe you want to add something here?
Another observation involves bugs that we could call {\it
severe} and they are a subset of the {\it Security} category.
Such bugs are related to vulnerabilities that appear due to the lack of user-input
validation and can lead to damaging attacks like {\sc sql} injection~\cite{RL12} and
Cross-Site Scripting~\cite{WS08}. Note that to exploit such vulnerabilities, a malicious user does
not have to know anything about the applications internals. For all the other
bugs, another program should be written either to incorporate references to
mutable objects, access non-final fields and others. In addition, as
bug descriptions indicate,\footnote{\url{http://findbugs.sourceforge.net/bugDescriptions.html}}
if an application has bugs like these, it might have more vulnerabilities that FindBugs doesn't
report. Table~\ref{tbl:sev} presents the number of {\sc jar}s where at least one of these
bugs exists. In the following section we provide results separately for this
subcategory which we call {\it Security High}. The remaining bugs of the {\it Security} category
are grouped together with the bugs of the {\it Malicious Code} category in another subcategory that
we call {\it Security Low}.

@@

By examining Maven as a whole we observed that security bugs and the
bugs coming from the {\it Bad Practice} category are the most popular types of
bugs in the ecosystem. This could be a strong indication that programmers write code
that implements the required functionality without considering its many
security aspects; an issue that has already been reported in
literature~\cite{SH09}. In addition, contrary to the Linus' Law, from the top fifty versions
(based on their PageRank), twenty five of them were the latest versions at the
time and at the same time contained security bugs.
This also highlights the {\it domino effect}~\cite{TH04}. No matter how well a programmer
secures a software component it won't matter if he or she is using vulnerable
code fragments from another library. Finally, 5,355 versions ($\approx 4,64\% $), contain at
least one severe security bug. Given the fact that other projects include these
versions as their dependencies, they are automatically rendered vulnerable if
they use the code fragments that include the defects.

>>>>

The main contributions of this research are:
\begin{itemize}
	\item The analysis of how security bugs evolve over time. To achieve
this, we inspect every project per version. We expect that security
bugs should decrease as a project evolves since they are critical and developers
should eliminate them.
	\item Security bug persistence between versions. We show evidence that
security bugs are not eliminated between versions.
	\item We show evidence of the relation between security bugs and a project
version's size.  Here, size is defined in terms of bytecode.
	\item The correlation of security bugs with other bug categories. We
show evidence on how such bugs are correlated with bugs that have to do with
performance, coding practices and others.
%	\item The relation of software dependencies and security bugs. Our
%hypothesis is that an artifact that is a dependency of many other artifacts
%should have a low rate of security bugs. 
\end{itemize}

@@

Figure~\ref{fig:bug-per} shows how software bugs are distributed among the
repository. Together with the {\it Bad Practice} bugs and the {\it Style} bugs,
security bugs (the sum of the {\it Security} and {\it Malicious Code}
categories) are the most popular in the repository ($\geq 22\%$ each).
% An interesting observation is that 4,908 projects did not
% have any bugs at all. These cases involved small artifacts with a minimum number of classes.
% This goes along with the relationship between the size of an artifact and it's
% bugs that is presented in the upcoming section.

% dimitro@dds: maybe you want to add something here?
Another observation involves bugs that we could call {\it
severe} and they are a subset of the {\it Security} category.
Such bugs are related to vulnerabilities that appear due to the lack of user-input
validation and can lead to damaging attacks like {\sc sql} injection~\cite{RL12} and
Cross-Site Scripting~\cite{WS08}. Note that to exploit such vulnerabilities, a malicious user does
not have to know anything about the applications internals. For all the other
bugs, another program should be written either to incorporate references to
mutable objects, access non-final fields and others. Table~\ref{tbl:sev} presents the number
of {\sc jar}s where at least one of these bugs exists. In the following section
we provide results separately for this subcategory which we call {\it Security High}.
The remaining bugs of the {\it Security} category
are grouped together with the bugs of the {\it Malicious Code} category
in another subcategory that we call {\it Security Low}.

@@

By examining Maven as a whole we observed that security bugs and the
bugs coming from the {\it Bad Practice} category are the most popular types of
bugs in the ecosystem. This could be a strong indication that programmers write code
that implements the required functionality without considering its many
security aspects; an issue that has already been reported in
literature~\cite{SH09}. In addition, contrary to the Linus' Law, from the top fifty versions
(based on their PageRank), twenty five of them were the latest versions at the
time and at the same time contained security bugs.
This also highlights the {\it domino effect}~\cite{TH04}. No matter how well a programmer
secures a software component it won't matter if he or she is using vulnerable
code fragments from another library. Finally, 5,355 versions ($\approx 4,64\% $), contained at
least one severe security bug. Given the fact that other projects include these
versions as their dependencies, they are automatically rendered vulnerable if
they use the code fragments that include the defects. Finally, as
bug descriptions indicate,\footnote{\url{http://findbugs.sourceforge.net/bugDescriptions.html}}
if an application has bugs coming from the {\it Security High} category,
it might have more vulnerabilities that FindBugs doesn't report.
